\pagestyle{empty}
\title{Optimal Parametric Search for Path and Tree Partitioning}

\author{Greg N. Frederickson\thanks{Supported in part by the National Science Foundation under
grants CCR-86202271 and CCR-9001241,
and by the Office of Naval Research under contract
N00014-86-K-0689.} \\
\and Samson Zhou\thanks{Research supported by NSF CCF-1649515.
Email: \email{samsonzhou@gmail.com}
}}
%\vspace{.15in}
%Department of Computer Science\\
%Purdue University\\
%West Lafayette, Indiana 47907\vspace{.1in}\\
%gnf@cs.purdue.edu
\date{\today}
\dspace
\maketitle
\pagestyle{plain}
\noindent
{\bf Abstract.}
We present linear-time algorithms for partitioning a path or a tree with weights on the vertices by removing $k$ edges to maximize the minimum-weight component. We also use the same framework to partition a path with weight on the vertices, removing $k$ edges to minimize the  maximum-weight component. The algorithms use the parametric search paradigm, testing candidate values until an optimum is found while simultaneously reducing the running time needed for each test.
For path-partitioning, the algorithm employs a synthetic weighting scheme that results in a constant fraction reduction in running time after each test. 
For tree-partitioning, our dual-pronged strategy makes progress no matter what the specific structure of our tree is.
\vspace{.2in}\\
{\bf Key words and phrases.}
Adaptive algorithms, 
data structures,
max-min,
min-max,
parametric search,
partial order,
path partitioning,
prune-and-search,
sorted matrices,
tree partitioning.
\newpage
\pagenumbering{arabic}
\dspace


